---
title: >
  Light Exposure Patterns under Permanent Daylight-Saving Time in Turkiye
author:
  - name: "Altug Didikoglu"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    theme: cosmo
    highlight: tango
    df_print: paged
    code_folding: hide
editor_options:
  chunk_output_type: console
---

# Overview

> This document presents a data analysis of the MeLiDos project (https://www.melidos.eu/). The project’s openly shared datasets are publicly available at https://github.com/MeLiDosProject. Türkiye (Izmir) dataset, which includes personal light exposure measurements collected using head-mounted devices and wristbands, is available at https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025. The data, preprocessing code, and variable names used in this analysis correspond to the repository versions available as of 16 July 2026. An additional personal light exposure dataset was collected using the same light-logging device, ActLumus Condor wristband. For this newly collected dataset, only the cleaned data corresponding to the version available on 16 July 2026 were included in the analysis.

# Project setup

```{r, message=FALSE, warning=FALSE}

# Clean environment
rm(list = ls())

# Knit options
options(tibble.print_max = Inf)

# List of required packages
packages <- c(
  "LightLogR","tidyverse","dplyr","lubridate","suncalc","ggplot2","patchwork",
  "sf","rnaturalearth","rnaturalearthdata","purrr",
  "melidosData","gtsummary","gt","lme4","lmerTest", "emmeans","tibble",
  "broom.mixed","performance","effectsize","data.table"
)

# Download packages if required 
packages[!packages %in% rownames(installed.packages())] |>
  install.packages(dependencies = TRUE)

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# R version
R.version.string

# Loaded packages and versions
data.frame(
  package = names(sessionInfo()$otherPkgs[packages]),
  version = sapply(sessionInfo()$otherPkgs[packages], function(x) x$Version),
  row.names = NULL
)
rm(packages)

```

# Data

```{r, message=FALSE, warning=FALSE}

# Location info
tz <-  "Europe/Istanbul"
coordinates <- c(38.3, 26.6)
lat <- coordinates[1]
lon <- coordinates[2]

# Import data
sleepdiary    <- load_data("sleepdiaries", site = "IZTECH")
current       <- load_data("currentconditions", site = "IZTECH")
demographics  <- load_data("demographics", site = "IZTECH")
# light  <- load_data("light_wrist", site = "IZTECH")
# light_glasses  <- load_data("light_glasses", site = "IZTECH")
load_rdata <- function(path) {
  env <- new.env()
  object_names <- load(path, envir = env)
  env[[object_names]]
}
sleepdiary2   <- load_rdata("data/sleep_diary_iztech.RData")
demographics2 <- load_rdata("data/demographics_iztech.RData")
# light2        <- load_rdata("data/light_wrist_iztech.RData")

# Bind data
# light_all <- bind_rows(
#     light  |> select(Id, Datetime, MEDI, LIGHT),
#     light2 |> select(Id, Datetime, MEDI, LIGHT)
#   ) |>
#   arrange(Id, Datetime)
demographics_all <- bind_rows(
    demographics  |> select(Id, age, sex),
    demographics2 |> select(Id, age, sex)
  ) |>
  mutate(
    sex = case_when(
      tolower(trimws(as.character(sex))) == "female" ~ "Female",
      tolower(trimws(as.character(sex))) == "male"   ~ "Male",
      TRUE                                           ~ NA_character_
    )
  )
sleepdiary_all <- bind_rows(
  sleepdiary |>
    transmute(
      Id,
      bedtime = sleepprep,
      waketime = wake,
      sleepdelay,
      awakenings,
      awake_duration,
      sleepquality = as.numeric(sleepquality),
      day_type_today = daytype2,
      sleep_duration = as.numeric(sleep_duration)
    ),
  sleepdiary2 |>
    filter(sleep_type == "main_sleep") |>
    transmute(
      Id,
      bedtime = datetime_sleep_attempt,
      waketime = datetime_wake,
      sleepdelay = sleep_latency_min,
      awakenings,
      awake_duration = awake_duration_min,
      sleepquality = sleep_quality,
      day_type_today = day_type,
      sleep_duration = as.numeric(sleep_duration_min) / 60
    )
  ) |>
  mutate(
    day_type_today = case_when(
      tolower(trimws(as.character(day_type_today))) == "a free day" ~ "free day",
      tolower(trimws(as.character(day_type_today))) == "free_day" ~ "free day",
      tolower(trimws(as.character(day_type_today))) == "a work day"   ~ "work day",
      tolower(trimws(as.character(day_type_today))) == "workday"   ~ "work day",
      TRUE ~ NA_character_
    )
  )|>
  arrange(Id, waketime)

# Clean light data
# light_all <- light_all |>
#   mutate(
#     MEDI = replace(
#       as.numeric(MEDI),
#       MEDI >= 120000,
#       NA_real_
#     )
#   ) |>
#   cut_Datetime(
#     unit = "1 hour",
#     group_by = TRUE,
#     type = "floor"
#   ) |>
#   remove_partial_data(
#     MEDI,
#     threshold.missing = 0.5
#   ) |>
#   ungroup(Datetime.rounded) |>
#   select(-Datetime.rounded) |>
#   add_Date_col(group.by = TRUE) |>
#   gap_handler(full.days = TRUE) |>
#   remove_partial_data(
#     MEDI,
#     threshold.missing = 0.2
#   ) |>
#   ungroup(Date) |>
#   add_photoperiod(
#     coordinates = coordinates,
#     Datetime.colname = Datetime
#   ) |>
#   arrange(Id, Datetime)
# save(light_all, file = "data/light_all.Rdata")
load("data/light_all.Rdata")

# Clean light glasses data
# light_glasses <- light_glasses |>
#   mutate(
#     MEDI = replace(
#       as.numeric(MEDI),
#       MEDI >= 120000,
#       NA_real_
#     )
#   ) |>
#   cut_Datetime(
#     unit = "1 hour",
#     group_by = TRUE,
#     type = "floor"
#   ) |>
#   remove_partial_data(
#     MEDI,
#     threshold.missing = 0.5
#   ) |>
#   ungroup(Datetime.rounded) |>
#   select(-Datetime.rounded) |>
#   add_Date_col(group.by = TRUE) |>
#   gap_handler(full.days = TRUE) |>
#   remove_partial_data(
#     MEDI,
#     threshold.missing = 0.2
#   ) |>
#   ungroup(Date) |>
#   add_photoperiod(
#     coordinates = coordinates,
#     Datetime.colname = Datetime
#   ) |>
#   arrange(Id, Datetime)
# save(light_glasses, file = "data/light_glasses.Rdata")
load("data/light_glasses.Rdata")

# Calculate light metrics
# metric_light <-
#   light_all |> 
#   group_by(Id, Date) |> 
#   summarize(
#     Mean = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
#     dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
#     MDER = mean(MEDI / LIGHT, na.rm = TRUE),
#     duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     timing_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     timing_above_threshold(MEDI, Datetime, "above", 10, na.rm = TRUE, as.df = TRUE),
#     period_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     .groups = "drop_last"
#   ) |> 
#   mutate(MDER = ifelse(abs(MDER) == Inf, NA, MDER))
# metric_light_morning <-
#   light_all |> 
#   filter(hour(Datetime) >= 6 & hour(Datetime) <= 9) |>
#   group_by(Id, Date) |> 
#   summarize(
#     Mean_Morning = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
#     dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
#     MDER_Morning = mean(MEDI / LIGHT, na.rm = TRUE),
#     duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     .groups = "drop_last"
#   ) |>
#   rename(duration_above_250_Morning = duration_above_250) |>
#   rename(dose_Morning = dose) |> 
#   mutate(MDER_Morning = ifelse(abs(MDER_Morning) == Inf, NA, MDER_Morning))
# light_metrics <-
#   metric_light |> 
#   left_join(metric_light_morning, 
#             by = c("Id", "Date")) |>
#   Datetime2Time() |>
#   mutate(
#     across(c(where(hms::is_hms), where(is.duration)), 
#            \(x) as.numeric(x) / 60 / 60)
#   ) |> 
#   filter(Mean > 0)
# save(light_metrics, file = "data/light_metrics.Rdata")
load("data/light_metrics.Rdata")

# Calculate light glasses metrics
# metric_light_glasses <-
#   light_glasses |> 
#   group_by(Id, Date) |> 
#   summarize(
#     Mean = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
#     dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
#     MDER = mean(MEDI / LIGHT, na.rm = TRUE),
#     duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     timing_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     timing_above_threshold(MEDI, Datetime, "above", 10, na.rm = TRUE, as.df = TRUE),
#     period_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     .groups = "drop_last"
#   ) |> 
#   mutate(MDER = ifelse(abs(MDER) == Inf, NA, MDER))
# metric_light_glasses_morning <-
#   light_glasses |> 
#   filter(hour(Datetime) >= 6 & hour(Datetime) <= 9) |>
#   group_by(Id, Date) |> 
#   summarize(
#     Mean_Morning = MEDI |> log_zero_inflated() |> mean(na.rm = TRUE) |> exp_zero_inflated(),
#     dose(MEDI, Datetime, na.rm = TRUE, as.df = TRUE),
#     MDER_Morning = mean(MEDI / LIGHT, na.rm = TRUE),
#     duration_above_threshold(MEDI, Datetime, "above", 250, na.rm = TRUE, as.df = TRUE),
#     .groups = "drop_last"
#   ) |>
#   rename(duration_above_250_Morning = duration_above_250) |>
#   rename(dose_Morning = dose) |> 
#   mutate(MDER_Morning = ifelse(abs(MDER_Morning) == Inf, NA, MDER_Morning))
# light_metrics_glasses <-
#   metric_light_glasses |> 
#   left_join(metric_light_glasses_morning, 
#             by = c("Id", "Date")) |>
#   Datetime2Time() |>
#   mutate(
#     across(c(where(hms::is_hms), where(is.duration)), 
#            \(x) as.numeric(x) / 60 / 60)
#   ) |> 
#   filter(Mean > 0)
# save(light_metrics_glasses, file = "data/light_metrics_glasses.Rdata")
load("data/light_metrics_glasses.Rdata")

# Define the ID lists (pre and post Spring clock change)
id_periods <- light_all |>
  mutate(
    Date  = as.Date(Datetime),
    month = month(Date),
    day   = day(Date)
  ) |>
  group_by(Id) |>
  summarise(
    has_post_period = all(
      (month > 4 & month < 10) |
        (month == 4 & day >= 1) |
        (month == 10 & day <= 25),
      na.rm = TRUE
    ),
    has_pre_period = all(
      month == 11 |
        month == 12 |
        month == 1 |
        month == 2 |
        (month == 3 & day <= 25),
      na.rm = TRUE
    ),
    .groups = "drop"
  )
post_ids <- id_periods |>
  filter(has_post_period) |>
  pull(Id)
pre_ids <- id_periods |>
  filter(has_pre_period) |>
  pull(Id)

# Add before/after probable DST spring change date (End of March)
light_all <- light_all |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
light_metrics <- light_metrics |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
light_glasses <- light_glasses |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
light_metrics_glasses <- light_metrics_glasses |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
current <- current |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
sleepdiary_all <- sleepdiary_all |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))
demographics_all <- demographics_all |> 
  filter(Id %in% c(pre_ids, post_ids)) |> 
  mutate(season = case_when(
    Id %in% pre_ids  ~ "November-March",
    Id %in% post_ids ~ "April-October"
  ))

# Keep required objects
keep_objects <- c(
  "coordinates",
  "current",
  "demographics_all",
  "lat",
  "light_all",
  "light_glasses",
  "light_metrics",
  "light_metrics_glasses",
  "lon",
  "sleepdiary_all",
  "tz"
)
rm(list = setdiff(ls(), keep_objects))

```

# Solar cycle models

```{r, message=FALSE, warning=FALSE}

# Permanent DST (2025) graph - Izmir
dates <- seq.Date(from = as.Date("2025-01-01"),
                  to   = as.Date("2025-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = lat,
  lon = lon,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) + minute(dawn)/60,
    dusk_hour = hour(dusk) + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (A) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
A <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2025-01-01"), to = as.Date("2025-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Izmir-Permanent DST (UTC+3)", y = "Hour of day") +
  theme_minimal() 

# Biannual shifting (2015) graph - Izmir
dates <- seq.Date(from = as.Date("2015-01-01"),
                  to   = as.Date("2015-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = lat,
  lon = lon,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) + minute(dawn)/60,
    dusk_hour = hour(dusk) + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (B) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
B <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2015-01-01"), to = as.Date("2015-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Izmir-Biannual Shifting", y = "Hour of day") +
  theme_minimal()  

# Permanent SDT (2025) graph - Izmir
dates <- seq.Date(from = as.Date("2025-01-01"),
                  to   = as.Date("2025-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = lat,
  lon = lon,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) -1 + minute(dawn)/60,
    dusk_hour = hour(dusk) -1 + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (C) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
C <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2025-01-01"), to = as.Date("2025-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Izmir-Permanent SDT (UTC+2)", y = "Hour of day") +
  theme_minimal()   

# Permanent DST (2025) graph - Van
dates <- seq.Date(from = as.Date("2025-01-01"),
                  to   = as.Date("2025-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = 38.3,
  lon = 43.3,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) + minute(dawn)/60,
    dusk_hour = hour(dusk) + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (D) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
D <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2025-01-01"), to = as.Date("2025-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Van-Permanent DST (UTC+3)", y = "Hour of day") +
  theme_minimal() 

# Biannual shifting (2015) graph - Van
dates <- seq.Date(from = as.Date("2015-01-01"),
                  to   = as.Date("2015-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = 38.3,
  lon = 43.3,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) + minute(dawn)/60,
    dusk_hour = hour(dusk) + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (E) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
E <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2015-01-01"), to = as.Date("2015-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Van-Biannual Shifting", y = "Hour of day") +
  theme_minimal()   

# Permanent SDT (2025) graph - Van
dates <- seq.Date(from = as.Date("2025-01-01"),
                  to   = as.Date("2025-12-31"),
                  by   = "day")
sun_times <- getSunlightTimes(
  date = dates,
  lat = 38.3,
  lon = 43.3,
  keep = c("dawn", "dusk"),
  tz = tz
)
sun_times <- sun_times %>%
  mutate(
    dawn_hour = hour(dawn) -1 + minute(dawn)/60,
    dusk_hour = hour(dusk) -1 + minute(dusk)/60
  )
red_band <- data.frame(
  xmin = min(sun_times$date) - 1,
  xmax = max(sun_times$date) + 1,
  ymin = 6,
  ymax = 9
)
# Probability of dawn after 6am
cat("===== (F) Probability of dawn after 6am:")
sun_prob <- sun_times |>
  mutate(
    start = 6,
    end = 9,
    after_dawn = pmax(0, end - pmax(dawn_hour, start)),
    before_dawn = pmax(0, pmin(dawn_hour, end) - start),
    prob_after_dawn  = after_dawn / (end - start),
    prob_before_dawn = before_dawn / (end - start)
  ) |>
  summarise(
    mean_prob_after_dawn  = mean(prob_after_dawn),
    mean_prob_before_dawn = mean(prob_before_dawn)
  )
sun_prob
# Plot
F <- ggplot(sun_times) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = 0, ymax = dawn_hour),
            fill = "black", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dawn_hour, ymax = dusk_hour),
            fill = "orange", alpha = 0.5) +
  geom_rect(aes(xmin = date - 0.5, xmax = date + 0.5, ymin = dusk_hour, ymax = 24),
            fill = "black", alpha = 0.5) +
  geom_rect(data = red_band,
            aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
            fill = "red", alpha = 0.2) +
  scale_x_date(
    breaks = seq.Date(from = as.Date("2025-01-01"), to = as.Date("2025-12-01"), by = "month"),
    labels = function(x) substr(format(x, "%b"), 1, 1)
  ) +
  scale_y_continuous(
    breaks = seq(0, 24, by = 4),
    limits = c(0, 24)
  ) +
  labs(x = "Van-Permanent SDT (UTC+2)", y = "Hour of day") +
  theme_minimal()   

# Graph panel
P_panel <- A + B + C + D + E + F +
  plot_layout(ncol = 3, guides = "collect") & 
  plot_annotation(tag_levels = 'A') & 
  theme(
    axis.title.y = element_text(size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )
P_panel

# Save
ggsave(
  filename = "Figure-Dawn to dusk.png", 
  plot = P_panel,
  width = 30,
  height = 25,
  units = "cm",
  dpi = 500
)

```

# Study descriptions

```{r}

# Map plot
world <- ne_countries(scale = "medium", returnclass = "sf")
A <- ggplot(data = world) +
  geom_sf(color = "black", size = 0.3) +
  geom_sf(data = subset(world, admin == "Turkey"), fill = "#ffe5d9", color = "black", size = 0.3) +
  annotate("point", x = 27.14, y = 38.42, color = "red", size = 2) +
  annotate("text", x = 33.30, y = 38.52, label = "Izmir, Turkiye", color = "black", size = 3) +
  geom_vline(xintercept = c(0, 15, 30, 45, 60),  color = "blue", linewidth = 0.3) +
  coord_sf(xlim = c(-10, 65), ylim = c(20, 60), expand = FALSE) +
  scale_x_continuous(breaks = c(0, 15, 30, 45, 60)) +
  scale_y_continuous(breaks = seq(20, 60, by = 10)) +
  guides(fill = guide_legend(nrow = 1, title.position = "top", title.hjust = 0.5)) +
  theme_minimal() +
  labs(x = "Longitude", y = "Latitude") +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )

# Study participant dates plot
major_breaks <- as.POSIXct(
  c(
    "2024-12-31",
    "2025-03-31",
    "2025-10-31",
    "2026-03-31"
  ),
  tz = "Europe/Istanbul"
)
monthly_breaks <- seq(
  from = as.POSIXct("2024-12-01", tz = "Europe/Istanbul"),
  to   = as.POSIXct("2026-03-01", tz = "Europe/Istanbul"),
  by   = "1 month"
)
light_all_plot <- light_all |>
  mutate(
    participant_group = factor(
      if_else(
        startsWith(as.character(Id), "IZTECH"),
        "IZTECH",
        "Other"
      ),
      levels = c("Other", "IZTECH")
    )
  ) |>
  group_by(participant_group)
B <- light_all_plot |>
  gg_overview(color = participant_group) +
  scale_color_manual(
    values = c(
      "IZTECH" = "#D55E00",
      "Other"  = "#4D4D4D"
    ),
    name = "Participant group"
  ) +
  scale_x_datetime(
    breaks = major_breaks,
    labels = scales::label_date("%Y-%m"),
    minor_breaks = monthly_breaks
  ) +
  geom_vline(
    xintercept = as.POSIXct(
      c(
        "2025-03-31 00:00:00",
        "2025-10-31 00:00:00",
        "2026-03-31 00:00:00"
      ),
      tz = "Europe/Istanbul"
    ),
    linetype = "dashed",
    linewidth = 0.6,
    color = "black"
  ) +
  labs(
    x = NULL,
    y = "Participants"
  ) +
  theme(
    axis.text.y = element_blank(),
    panel.grid.minor.x = element_line(
      color = "grey88",
      linewidth = 0.3
    ),
    legend.position = "none"
  )

# Graph panel
P_panel <- (A | B) +
  plot_layout(widths = c(1, 1.3)) +
  plot_annotation(tag_levels = "A") &
  theme(
    axis.title.y = element_text(size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )
P_panel

# Save
ggsave(
  filename = "Figure-Map and participant.png", 
  plot = P_panel,
  width = 30,
  height = 15,
  units = "cm",
  dpi = 300
)

# Demographics summary
demographics_summary <- demographics_all |>
  summarise(
    n = n(),
    n_melidos = sum(startsWith(as.character(Id), "IZTECH"), na.rm = TRUE),
    n_AprilToOctober = sum(season == "April-October", na.rm = TRUE),
    n_melidos_AprilToOctober = sum(startsWith(as.character(Id), "IZTECH") & 
                                     season == "April-October", na.rm = TRUE),
    age_mean = mean(age, na.rm = TRUE),
    age_sd = sd(age, na.rm = TRUE),
    female_n = sum(sex == "Female", na.rm = TRUE),
    male_n = sum(sex == "Male", na.rm = TRUE),
    male_percent = 100 * mean(sex == "Male", na.rm = TRUE),
    female_percent = 100 * mean(sex == "Female", na.rm = TRUE),    
  )
demographics_summary

# Summary light measures valid data per participant
summary(
  light_metrics |>
  group_by(Id) |>
  summarise(
    n_row = n(),
    .groups = "drop"
  ) |>
  arrange(desc(n_row)) |>
  select(n_row)
)

# Valid light days
nrow(light_metrics)
nrow(light_metrics_glasses)

# Summary sleep measures valid data per participant
summary(
  sleepdiary_all |>
  group_by(Id) |>
  summarise(
    n_row = n(),
    .groups = "drop"
  ) |>
  arrange(desc(n_row)) |>
  select(n_row)
)

# Valid sleep days
nrow(sleepdiary_all)

# Summary ema measures valid data per participant
summary(
  current |>
  group_by(Id) |>
  summarise(
    n_row = n(),
    .groups = "drop"
  ) |>
  arrange(desc(n_row)) |>
  select(n_row)
)

# Valid ema days
nrow(current)

# Description of dusk/dawn
light_all |>
  ungroup() |>
  summarise(
    dawn_time_mean = mean(hour(dawn) + minute(dawn) / 60, na.rm = TRUE),
    dusk_time_mean = mean(hour(dusk) + minute(dusk) / 60, na.rm = TRUE),
    dawn_time_sd   = sd(hour(dawn) + minute(dawn) / 60, na.rm = TRUE),
    dusk_time_sd   = sd(hour(dusk) + minute(dusk) / 60, na.rm = TRUE),

    photoperiod_mean = mean(
      as.numeric(difftime(dusk, dawn, units = "hours")),
      na.rm = TRUE
    ),
    photoperiod_min = min(
      as.numeric(difftime(dusk, dawn, units = "hours")),
      na.rm = TRUE
    ),
    photoperiod_max = max(
      as.numeric(difftime(dusk, dawn, units = "hours")),
      na.rm = TRUE
    )
  )
light_all |>
  group_by(season) |>
  summarise(
    dawn_time_mean = mean(hour(dawn) + minute(dawn)/60, na.rm = TRUE),
    dusk_time_mean = mean(hour(dusk) + minute(dusk)/60, na.rm = TRUE),
    dawn_time_sd = sd(hour(dawn) + minute(dawn)/60, na.rm = TRUE),
    dusk_time_sd = sd(hour(dusk) + minute(dusk)/60, na.rm = TRUE),
    .groups = "drop"
  )

```

# Light metrics description

```{r, message=FALSE, warning=FALSE}

# Light metrics description table
light_metrics_table <- 
  light_metrics |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) |>
  select(
    Id, Date,
    Mean, dose, MDER,
    mean_timing_above_250,
    first_timing_above_250,
    last_timing_above_250,
    duration_above_250,
    period_above_250,
    Mean_Morning,
    dose_Morning,
    MDER_Morning,
    duration_above_250_Morning,
    season
  ) |>
  group_by(season) |>
  select(-Id, -Date) |>
  tbl_summary(by = season,
              statistic = list(all_continuous() ~ "{mean} (±{sd})"),
              missing = "no",
              label = list(
                mean_timing_above_250 = 
                  "mean timing above 250 lx mel EDI",
                first_timing_above_250 = 
                  "first time above 250 lx mel EDI",
                last_timing_above_250 = 
                  "last time above 250 lx mel EDI",
                duration_above_250 = "duration above 250 lx mel EDI",
                period_above_250 = "length of longest continuous period above 250 lx mel EDI",
                Mean = "average mel EDI (lx)",
                dose = "melanopic light exposure (lx·h)",
                MDER = "melanopic daylight efficacy ratio",
                Mean_Morning = "average mel EDI (lx) between 06:00 - 09:00",
                dose_Morning = "melanopic light exposure (lx·h) between 06:00 - 09:00",
                MDER_Morning = "melanopic daylight efficacy ratio between 06:00 - 09:00",
                duration_above_250_Morning = "duration above 250 lx mel EDI between 06:00 - 09:00"
              )
  )
light_metrics_table

# Save
gt_tbl <- as_gt(light_metrics_table)
options(
  chromote.timeout = 60,
  chromote.headless = "new"
)
gtsave(gt_tbl, "Table-Light metrics wrist description.png", vwidth = 1200)
gtsave(gt_tbl, "Table-Light metrics wrist description.docx", vwidth = 1200)

# Light metrics description table
light_metrics_glasses_table <- 
  light_metrics_glasses |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) |>
  select(
    Id, Date,
    Mean, dose, MDER,
    mean_timing_above_250,
    first_timing_above_250,
    last_timing_above_250,
    duration_above_250,
    period_above_250,
    Mean_Morning,
    dose_Morning,
    MDER_Morning,
    duration_above_250_Morning,
    season
  ) |>
  group_by(season) |>
  select(-Id, -Date) |>
  tbl_summary(by = season,
              statistic = list(all_continuous() ~ "{mean} (±{sd})"),
              missing = "no",
              label = list(
                mean_timing_above_250 = 
                  "mean timing above 250 lx mel EDI",
                first_timing_above_250 = 
                  "first time above 250 lx mel EDI",
                last_timing_above_250 = 
                  "last time above 250 lx mel EDI",
                duration_above_250 = "duration above 250 lx mel EDI",
                period_above_250 = "length of longest continuous period above 250 lx mel EDI",
                Mean = "average mel EDI (lx)",
                dose = "melanopic light exposure (lx·h)",
                MDER = "melanopic daylight efficacy ratio",
                Mean_Morning = "average mel EDI (lx) between 06:00 - 09:00",
                dose_Morning = "melanopic light exposure (lx·h) between 06:00 - 09:00",
                MDER_Morning = "melanopic daylight efficacy ratio between 06:00 - 09:00",
                duration_above_250_Morning = "duration above 250 lx mel EDI between 06:00 - 09:00"
              )
  )
light_metrics_glasses_table

# Save
gt_tbl <- as_gt(light_metrics_glasses_table)
gtsave(gt_tbl, "Table-Light metrics glasses description.png", vwidth = 1200)
gtsave(gt_tbl, "Table-Light metrics glasses description.docx", vwidth = 1200)

```

# Light metrics compare

```{r, message=FALSE, warning=FALSE}

# Variable labels
labels <- c(
                mean_timing_above_250 = 
                  "mean timing above 250 lx mel EDI",
                first_timing_above_250 = 
                  "first time above 250 lx mel EDI",
                last_timing_above_250 = 
                  "last time above 250 lx mel EDI",
                duration_above_250 = "duration above 250 lx mel EDI",
                period_above_250 = "length of longest continuous period above 250 lx mel EDI",
                mean_timing_above_10 = 
                  "mean timing above 10 lx mel EDI",
                first_timing_above_10 = 
                  "first time above 10 lx mel EDI",
                last_timing_above_10 = 
                  "last time above 10 lx mel EDI",
                Mean = "average mel EDI (lx)",
                dose = "melanopic light exposure (lx·h)",
                MDER = "melanopic daylight efficacy ratio",
                Mean_Morning = "average mel EDI (lx) (morning)",
                dose_Morning = "melanopic light exposure (lx·h) (morning)",
                MDER_Morning = "melanopic daylight efficacy ratio (morning)",
                duration_above_250_Morning = "duration above 250 lx mel EDI (morning)",
  season = "months"
)

# Outcomes to analyse
outcomes <- c(
  "Mean",
  "Mean_Morning",
  "dose",
  "dose_Morning",
  "MDER",
  "MDER_Morning",
  "period_above_250",
  "mean_timing_above_250",
  "first_timing_above_250",
  "last_timing_above_250",
  "duration_above_250",
  "duration_above_250_Morning"
)

# Main LMM function
extract_lmm_results <- function(data, outcomes, predictor = "season", id_var = "Id") {
  
  purrr::map_dfr(outcomes, function(outcome) {
    
    formula_obj <- as.formula(
      paste0(outcome, " ~ ", predictor, " + (1 | ", id_var, ")")
    )
    
    model <- lmerTest::lmer(formula_obj, data = data, REML = TRUE)
    
    # ANOVA table
    aov_tab <- anova(model) |>
      as.data.frame() |>
      tibble::rownames_to_column("predictor_term")
    
    pred_row <- aov_tab[aov_tab$predictor_term == predictor, , drop = FALSE]
    
    # Random-effect SDs
    vc <- as.data.frame(VarCorr(model))
    random_sd_id <- vc$sdcor[vc$grp == id_var][1]
    random_sd_residual <- vc$sdcor[vc$grp == "Residual"][1]
    
    # ICC
    icc_val <- tryCatch(
      performance::icc(model)$ICC_adjusted,
      error = function(e) NA_real_
    )
    
    # Partial eta squared
    partial_eta2_val <- tryCatch({
      eta_tab <- effectsize::eta_squared(model, partial = TRUE) |>
        as.data.frame()
      eta_tab$Eta2_partial[eta_tab$Parameter == predictor][1]
    }, error = function(e) NA_real_)
    
    # R2
    r2_vals <- tryCatch(
      performance::r2_nakagawa(model),
      error = function(e) NULL
    )
    
    r2_marginal <- if (!is.null(r2_vals)) r2_vals$R2_marginal else NA_real_
    r2_conditional <- if (!is.null(r2_vals)) r2_vals$R2_conditional else NA_real_
    
    tibble::tibble(
      outcome = outcome,
      predictor = predictor,
      df1 = pred_row$NumDF[1],
      df2 = pred_row$DenDF[1],
      F_value = pred_row$`F value`[1],
      p_value = pred_row$`Pr(>F)`[1],
      n_obs = nobs(model),
      n_group = dplyr::n_distinct(data[[id_var]]),
      random_sd_id = random_sd_id,
      random_sd_residual = random_sd_residual,
      var_explained_id = icc_val,
      partial_eta2 = partial_eta2_val,
      r2_marginal = r2_marginal,
      r2_conditional = r2_conditional
    )
  })
}

# Run all models
light_metrics <- light_metrics |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  )
results_lmm <- extract_lmm_results(
  data = light_metrics,
  outcomes = outcomes,
  predictor = "season",
  id_var = "Id"
)

# Std. coefficient function
get_std_beta <- function(model, predictor) {
  std_pars <- parameters::standardize_parameters(model) |>
    as.data.frame()
  
  idx <- grep(predictor, std_pars$Parameter)
  std_pred <- std_pars[idx, , drop = FALSE]
  
  if (nrow(std_pred) == 1) {
    return(unname(as.numeric(std_pred$Std_Coefficient[1])))
  } else {
    return(NA_real_)
  }
}

# Add standardized beta manually with another plain loop
std_beta <- numeric(length(outcomes))
for (i in seq_along(outcomes)) {
  outcome <- outcomes[i]
  model <- lmerTest::lmer(
    as.formula(paste0(outcome, " ~ season + (1 | Id)")),
    data = light_metrics,
    REML = TRUE
  )
  std_beta[i] <- get_std_beta(model, "season")
}
results_lmm$standardised_beta_coefficient <- std_beta

# FDR adjusted p-value
results_lmm$p_value <- p.adjust(results_lmm$p_value, method = "fdr")

# Result table order
results_lmm <- results_lmm[, c(
  "outcome",
  "standardised_beta_coefficient",
  "n_obs",
  "n_group",
  "random_sd_id",
  "random_sd_residual",
  "var_explained_id",
  "partial_eta2",
  "r2_marginal",
  "r2_conditional",
  "p_value"
)]

# Results table gt
table_lmm_results <- function(results_object, title, subtitle, labels) {
  results_object |>
    arrange(p_value) |>
    mutate(
      outcome = recode(outcome, !!!labels)
    ) |>
    gt() |>
    fmt_number(
      columns = c(n_obs, n_group),
      decimals = 1
    ) |>
    fmt_number(
      columns = c(
        standardised_beta_coefficient,
        random_sd_id,
        random_sd_residual,
        var_explained_id,
        partial_eta2,
        r2_marginal,
        r2_conditional
      ),
      decimals = 3
    ) |>
    fmt_scientific(
      columns = p_value,
      decimals = 2
    ) |>
    cols_label(
      outcome = "Outcome",
      standardised_beta_coefficient = "Standardised beta",
      n_obs = "N observations",
      n_group = "N participants",
      random_sd_id = "SD (Id)",
      random_sd_residual = "SD (Residual)",
      var_explained_id = "ICC (Id)",
      partial_eta2 = "Partial Eta²",
      r2_marginal = "R² marginal",
      r2_conditional = "R² conditional",
      p_value = "p value"
    ) |>
    cols_align(
      align = "center",
      columns = everything()
    ) |>
    tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    tab_footnote(
      footnote = "Morning exposure values are calculated between 6 and 9 a.m.",
      locations = cells_column_labels(columns = outcome)
    ) |>
    tab_footnote(
      footnote = "Reference category: November - March",
      locations = cells_column_labels(columns = standardised_beta_coefficient)
    ) |>
    tab_footnote(
      footnote = "Ligth exposure metrics [wrist] ~ Time policy + (1 | Id).",
      locations = cells_title(groups = "subtitle")
    ) |>
    tab_footnote(
      footnote = "p values are adjusted for multiple testing using the false discovery rate (FDR) method.",
      locations = cells_column_labels(columns = p_value)
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(rows = p_value < 0.05)
    ) |>
    tab_options(
      table.font.size = "small"
    )
}

gt_lmm_results <- table_lmm_results(
  results_object = results_lmm,
  title = md(
    "**Comparison of personal light exposure between November-March and April-October periods**"
  ),
  subtitle = "Linear mixed model results",
  labels = labels
)

gt_lmm_results

# Save
gt_lmm_results <- gt_lmm_results |>
  gt::tab_options(
    table.font.size = 8,
    data_row.padding = gt::px(2),
    heading.padding = gt::px(2)
  )
gtsave(gt_lmm_results, "Table-Light metrics wrist comparisons.png", vwidth = 1200)
gtsave(gt_lmm_results, "Table-Light metrics wrist comparisons.docx", vwidth = 1200)

# Run all models for glasses
light_metrics_glasses <- light_metrics_glasses |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  )
results_lmm <- extract_lmm_results(
  data = light_metrics_glasses,
  outcomes = outcomes,
  predictor = "season",
  id_var = "Id"
)

# Add standardized beta manually with another plain loop
std_beta <- numeric(length(outcomes))
for (i in seq_along(outcomes)) {
  outcome <- outcomes[i]
  model <- lmerTest::lmer(
    as.formula(paste0(outcome, " ~ season + (1 | Id)")),
    data = light_metrics,
    REML = TRUE
  )
  std_beta[i] <- get_std_beta(model, "season")
}
results_lmm$standardised_beta_coefficient <- std_beta

# Result table order
results_lmm <- results_lmm[, c(
  "outcome",
  "standardised_beta_coefficient",
  "n_obs",
  "n_group",
  "random_sd_id",
  "random_sd_residual",
  "var_explained_id",
  "partial_eta2",
  "r2_marginal",
  "r2_conditional",
  "p_value"
)]

# Results table gt
table_lmm_results <- function(results_object, title, subtitle, labels) {
  results_object |>
    arrange(p_value) |>
    mutate(
      outcome = recode(outcome, !!!labels)
    ) |>
    gt() |>
    fmt_number(
      columns = c(n_obs, n_group),
      decimals = 1
    ) |>
    fmt_number(
      columns = c(
        standardised_beta_coefficient,
        random_sd_id,
        random_sd_residual,
        var_explained_id,
        partial_eta2,
        r2_marginal,
        r2_conditional
      ),
      decimals = 3
    ) |>
    fmt_scientific(
      columns = p_value,
      decimals = 2
    ) |>
    cols_label(
      outcome = "Outcome",
      standardised_beta_coefficient = "Standardised beta",
      n_obs = "N observations",
      n_group = "N participants",
      random_sd_id = "SD (Id)",
      random_sd_residual = "SD (Residual)",
      var_explained_id = "ICC (Id)",
      partial_eta2 = "Partial Eta²",
      r2_marginal = "R² marginal",
      r2_conditional = "R² conditional",
      p_value = "p value"
    ) |>
    cols_align(
      align = "center",
      columns = everything()
    ) |>
    tab_header(
      title = title,
      subtitle = subtitle
    ) |>
    tab_footnote(
      footnote = "Morning exposure values are calculated between 6 and 9 a.m.",
      locations = cells_column_labels(columns = outcome)
    ) |>
    tab_footnote(
      footnote = "Reference category: November - March",
      locations = cells_column_labels(columns = standardised_beta_coefficient)
    ) |>
    tab_footnote(
      footnote = "Ligth exposure metrics [glasses] ~ Time policy + (1 | Id).",
      locations = cells_title(groups = "subtitle")
    ) |>
    tab_footnote(
      footnote = "p values are not adjusted for multiple testing.",
      locations = cells_column_labels(columns = p_value)
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = cells_body(rows = p_value < 0.05)
    ) |>
    tab_options(
      table.font.size = "small"
    )
}

gt_lmm_results <- table_lmm_results(
  results_object = results_lmm,
  title = md(
    "**Comparison of personal light exposure between November-March and April-October periods**"
  ),
  subtitle = "Linear mixed model results",
  labels = labels
)

gt_lmm_results

# Save
gt_lmm_results <- gt_lmm_results |>
  gt::tab_options(
    table.font.size = 8,
    data_row.padding = gt::px(2),
    heading.padding = gt::px(2)
  )
gtsave(gt_lmm_results, "Table-Light metrics glasses comparisons.png", vwidth = 1200)
gtsave(gt_lmm_results, "Table-Light metrics glasses comparisons.docx", vwidth = 1200)

```

# Light exposure plot

```{r, message=FALSE, warning=FALSE}

# Before March
A <-
  light_all[light_all$season=="November-March",] |> 
  ungroup() |> 
  select(Datetime, MEDI) |> 
  aggregate_Date(
    unit = "10 mins",
    numeric.handler = \(x) median(x, na.rm = TRUE),
    upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
    upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
    lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
    lower95= quantile(MEDI, 0.025, na.rm = TRUE)
  ) |> 
  add_photoperiod(coordinates) |> 
  gg_doubleplot(geom = "blank", 
                facetting = FALSE, 
                jco_col = FALSE,
                x.axis.label = "Local time (HH:MM), November - March",
                y.axis.label = "Melanopic EDI (lx)"
  ) |> 
  gg_photoperiod() +
  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#332288", alpha = 0.4) +
  geom_ribbon(aes(ymin = lower50, ymax = upper50), fill = "#332288", alpha = 0.4) +
  geom_line(aes(y = MEDI), linewidth = 1) +
  coord_cartesian(ylim = c(0, 100000)) +
  theme(plot.caption = ggtext::element_markdown())

# After March
B <-
  light_all[light_all$season=="April-October",] |> 
  ungroup() |> 
  select(Datetime, MEDI) |> 
  aggregate_Date(
    unit = "10 mins",
    numeric.handler = \(x) median(x, na.rm = TRUE),
    upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
    upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
    lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
    lower95= quantile(MEDI, 0.025, na.rm = TRUE)
  ) |> 
  add_photoperiod(coordinates) |> 
  gg_doubleplot(geom = "blank", 
                facetting = FALSE, 
                jco_col = FALSE,
                x.axis.label = "Local time (HH:MM), April - October",
                y.axis.label = "Melanopic EDI (lx)"
  ) |> 
  gg_photoperiod() +
  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#332288", alpha = 0.4) +
  geom_ribbon(aes(ymin = lower50, ymax = upper50), fill = "#332288", alpha = 0.4) +
  geom_line(aes(y = MEDI), linewidth = 1) +
  coord_cartesian(ylim = c(0, 100000)) +
  theme(plot.caption = ggtext::element_markdown())

# Morning dose
light_metrics_plot <- light_metrics |>
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  )
C <- ggplot(light_metrics_plot, aes(x = season, y = (dose_Morning+0.1))) +
  geom_boxplot(
    width = 0.55,
    outlier.shape = NA,
    color = "black",
    fill = "grey90",
    size = 0.5
  ) +
  geom_jitter(
    width = 0.12,
    size = 1.2,
    alpha = 0.5,
    color = "black"
  ) +
  scale_y_log10(
    breaks = c(0.1, 1, 10, 100, 1000, 10000, 100000),
    labels = c("0.1", "1", "10", "100", "1,000", "10,000", "100,000"),
    limits = c(0.09, 100000)
  ) +
  scale_x_discrete() +
  labs(
    x = "Months",
    y = "Melanopic light dose (lx·h)\n(06:00–09:00, log scale)"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.title = element_text(face = "plain"),
    axis.text = element_text(color = "black"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.4),
    plot.margin = margin(10, 10, 10, 10),
    plot.tag = element_text(face = "bold", size = 14)
  )

# MDER
D <- ggplot(light_metrics_plot, aes(x = season, y = MDER)) +
  geom_boxplot(
    width = 0.55,
    outlier.shape = NA,
    color = "black",
    fill = "grey90",
    size = 0.5
  ) +
  geom_jitter(
    width = 0.12,
    size = 1.2,
    alpha = 0.5,
    color = "black"
  ) +
  scale_y_continuous(
    breaks = seq(0.4,1.2,0.2),
    limits = c(0.4,1.2)
  ) +
  scale_x_discrete() +
  labs(
    x = "Months",
    y = "Melanopic daylight\nefficacy ratio"
  ) +
  theme_classic(base_size = 13) +
  theme(
    axis.title = element_text(face = "plain"),
    axis.text = element_text(color = "black"),
    axis.line = element_line(linewidth = 0.5),
    axis.ticks = element_line(linewidth = 0.4),
    plot.margin = margin(10, 10, 10, 10),
    plot.tag = element_text(face = "bold", size = 14)
  )

# Graph panel
P_panel <- (A / B / (C | D)) +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "A") &
  theme(
    axis.title.y = element_text(size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )
P_panel

# Save
ggsave(
  filename = "Figure-Light exposure wrist.png", 
  plot = P_panel,
  width = 20,
  height = 20,
  units = "cm",
  dpi = 300
)


# Before March
A <-
  light_glasses[light_glasses$season=="November-March",] |> 
  ungroup() |> 
  select(Datetime, MEDI) |> 
  aggregate_Date(
    unit = "10 mins",
    numeric.handler = \(x) median(x, na.rm = TRUE),
    upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
    upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
    lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
    lower95= quantile(MEDI, 0.025, na.rm = TRUE)
  ) |> 
  add_photoperiod(coordinates) |> 
  gg_doubleplot(geom = "blank", 
                facetting = FALSE, 
                jco_col = FALSE,
                x.axis.label = "Local time (HH:MM), November - March",
                y.axis.label = "Melanopic EDI (lx)"
  ) |> 
  gg_photoperiod() +
  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#332288", alpha = 0.4) +
  geom_ribbon(aes(ymin = lower50, ymax = upper50), fill = "#332288", alpha = 0.4) +
  geom_line(aes(y = MEDI), linewidth = 1) +
  coord_cartesian(ylim = c(0, 100000)) +
  theme(plot.caption = ggtext::element_markdown())

# After March
B <-
  light_glasses[light_glasses$season=="April-October",] |> 
  ungroup() |> 
  select(Datetime, MEDI) |> 
  aggregate_Date(
    unit = "10 mins",
    numeric.handler = \(x) median(x, na.rm = TRUE),
    upper95 = quantile(MEDI, 0.975, na.rm = TRUE),
    upper50 = quantile(MEDI, 0.75, na.rm = TRUE),
    lower50 = quantile(MEDI, 0.25, na.rm = TRUE),
    lower95= quantile(MEDI, 0.025, na.rm = TRUE)
  ) |> 
  add_photoperiod(coordinates) |> 
  gg_doubleplot(geom = "blank", 
                facetting = FALSE, 
                jco_col = FALSE,
                x.axis.label = "Local time (HH:MM), April - October",
                y.axis.label = "Melanopic EDI (lx)"
  ) |> 
  gg_photoperiod() +
  geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#332288", alpha = 0.4) +
  geom_ribbon(aes(ymin = lower50, ymax = upper50), fill = "#332288", alpha = 0.4) +
  geom_line(aes(y = MEDI), linewidth = 1) +
  coord_cartesian(ylim = c(0, 100000)) +
  theme(plot.caption = ggtext::element_markdown())

# Graph panel
P_panel <- (A / B) +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "A") &
  theme(
    axis.title.y = element_text(size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )
P_panel

# Save
ggsave(
  filename = "Figure-Light exposure glasses.png", 
  plot = P_panel,
  width = 20,
  height = 15,
  units = "cm",
  dpi = 300
)

```

# KSS-mood-sleep description

```{r, message=FALSE, warning=FALSE}

# Make momentary assessments numeric
current <- current |>
  mutate(across(c(kss, sad, elated, energetic, irritable, angry, anxious), as.numeric))

# Make sleep assessments numeric
sleep <- sleepdiary_all |>
  mutate(
    sleep_duration = as.numeric(sleep_duration, units = "hours"),
    bedtime_h = as.numeric(format(bedtime, "%H")) +
                  as.numeric(format(bedtime, "%M")) / 60 +
                  as.numeric(format(bedtime, "%S")) / 3600,
    waketime_h = as.numeric(format(waketime, "%H")) +
             as.numeric(format(waketime, "%M")) / 60 +
             as.numeric(format(waketime, "%S")) / 3600,
    bedtime_h = ifelse(bedtime_h < 12, bedtime_h, bedtime_h - 24),
    sleepquality_n = as.numeric(sleepquality)
  )

# Calculate current time of day
current$time <- hour(current$Datetime) + minute(current$Datetime)/60

# Morning KSS calculation
sleep_dt   <- copy(as.data.table(sleep))
current_dt <- copy(as.data.table(current))
sleep_dt[, waketime := as.POSIXct(waketime, tz = "Europe/Istanbul")]
current_dt[, Datetime := as.POSIXct(Datetime, tz = "Europe/Istanbul")]
sleep_dt[, row_id := .I]
current_lookup <- current_dt[
  ,
  .(
    Id,
    Datetime,
    kss_datetime = Datetime,
    kss_value = as.numeric(kss)
  )
]
setkey(current_lookup, Id, Datetime)
kss_matches <- current_lookup[
  sleep_dt,
  on = .(Id, Datetime = waketime),
  roll = "nearest",
  .(
    row_id = i.row_id,
    waketime = i.waketime,
    kss_datetime,
    kss_value
  )
]
kss_matches[
  ,
  time_difference_hours :=
    abs(as.numeric(difftime(
      kss_datetime,
      waketime,
      units = "hours"
    )))
]
kss_matches[
  ,
  kss_morning := fifelse(
    !is.na(time_difference_hours) &
      time_difference_hours < 2,
    kss_value,
    NA_real_
  )
]
sleep_dt[
  kss_matches,
  on = .(row_id),
  `:=`(
    kss_morning = i.kss_morning,
    kss_datetime = i.kss_datetime,
    kss_time_difference_hours = i.time_difference_hours
  )
]
sleep_dt[, row_id := NULL]
sleep <- as.data.frame(sleep_dt)
sleep <- sleep |>
  dplyr::select(
    -kss_datetime,
    -kss_time_difference_hours
  )

# Current descriptive table
current_table <- 
  current |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) |>
  select(
    kss, sad, elated, energetic, irritable, angry, anxious, season
  ) |>
  tbl_summary(
    by = season,
    type = list(
      c(kss, sad, elated, energetic, irritable, angry, anxious) ~ "continuous"
    ),
    statistic = list(
      c(kss, sad, elated, energetic, irritable, angry, anxious) ~ "{mean} (±{sd})"
    ),
    missing = "no",
    label = list(
      kss = "slepiness (KSS)",
      sad = "sad",
      elated = "elated",
      energetic = "energetic",
      irritable = "irritable",
      angry = "angry",
      anxious = "anxious"
    )
  ) |>
  bold_labels()
current_table

# Save
gt_tbl <- as_gt(current_table)
gtsave(gt_tbl, "Table-Momentary assessments description.png", vwidth = 1200)
gtsave(gt_tbl, "Table-Momentary assessments description.docx", vwidth = 1200)
# Sleep descriptive table
sleep_table <- 
  sleep |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) |>
  select(
    sleepdelay,
    awakenings,
    awake_duration,
    sleep_duration,
    bedtime_h,
    waketime_h,
    sleepquality_n,
    kss_morning,
    season
  ) |>
  tbl_summary(
    by = season,
    type = list(
      c(awakenings, sleepdelay, awake_duration, sleep_duration, bedtime_h, waketime_h, sleepquality_n, kss_morning) ~ "continuous"
    ),
    statistic = list(all_continuous() ~ "{mean} (±{sd})"),
    missing = "no",
    label = list(
      sleepdelay = "sleep onset latency (min)",
      awakenings = "number of awakenings",
      awake_duration = "wake after sleep onset (min)",
      sleep_duration = "sleep duration (h)",
      bedtime_h = "sleep time",
      waketime_h = "wake time",
      sleepquality_n = "sleep quality",
      kss_morning = "morning sleepiness (KSS)"
    )
  ) |>
  bold_labels()
sleep_table

# Save
gt_tbl <- as_gt(sleep_table)
gtsave(gt_tbl, "Table-Sleep description.png", vwidth = 1200)
gtsave(gt_tbl, "Table-Sleep description.docx", vwidth = 1200)

```

# KSS-mood-sleep-light compare

```{r, message=FALSE, warning=FALSE}

# Season factor
sleep <- sleep |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) 
current <- current |> 
  mutate(
    season = factor(season, levels = c("November-March", "April-October"))
  ) 

# Sleep assessments models
model = lmer(sleepdelay ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(awakenings ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(awake_duration ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(sleep_duration ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(bedtime_h ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(waketime_h ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(sleepquality_n ~ season + (1 | Id), data = sleep)
summary(model)
model = lmer(kss_morning ~ season + (1 | Id), data = sleep)
summary(model)

# Merge light metrics and sleep
light_metrics_sleep <- light_metrics |>
  mutate(Date = as.Date(Date)) |>
  left_join(
    sleep |>
      transmute(
        Id,
        Date = as.Date(waketime) - 1,
        bedtime_h,
        waketime_h,
        sleepdelay,
        awakenings,
        awake_duration,
        sleepquality_n,
        sleep_duration,
        kss_morning
      ),
    by = c("Id", "Date")
  )

# Sleep-light assessments models
model = lmer(sleepdelay ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(awakenings ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(awake_duration ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(sleep_duration ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(bedtime_h ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(waketime_h ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(sleepquality_n ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(kss_morning ~ duration_above_250 + (1 | Id), data = light_metrics_sleep)
summary(model)
anova(model)
eta_squared(model)
r2(model)


model = lmer(sleepdelay ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(awakenings ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(awake_duration ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(sleep_duration ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(bedtime_h ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(waketime_h ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(sleepquality_n ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(kss_morning ~ dose + (1 | Id), data = light_metrics_sleep)
summary(model)
anova(model)
eta_squared(model)
r2(model)


# Merge light metrics and sleep
light_metrics_sleep <- light_metrics |>
  mutate(Date = as.Date(Date)) |>
  left_join(
    sleep |>
      transmute(
        Id,
        Date = as.Date(waketime),
        kss_morning
      ),
    by = c("Id", "Date")
  )

# Sleep-light assessments models
model = lmer(kss_morning ~ dose_Morning + (1 | Id), data = light_metrics_sleep)
summary(model)
model = lmer(kss_morning ~ duration_above_250_Morning + (1 | Id), data = light_metrics_sleep)
summary(model)

# Momentary assessments models
model = lmer(kss ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(elated ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(anxious ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(sad ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(irritable ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(energetic ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)
model = lmer(angry ~ (time + I(time^2))*season + (1 | Id), data = current)
summary(model)

# Merge light metrics and current conditions
light_metrics_current <- current |>
  mutate(Date = as.Date(Datetime)) |>
  left_join(
    light_metrics,
    by = c("Id", "Date","season")
  )

# Momentary-light assessments models
model = lmer(kss ~ (time + I(time^2))*dose + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(elated ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)
anova(model)
eta_squared(model)
r2(model)
model = lmer(anxious ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(sad ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(irritable ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(energetic ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(angry ~ (time + I(time^2))*dose  + (1 | Id), data = light_metrics_current)
summary(model)

model = lmer(kss ~ (time + I(time^2))*duration_above_250 + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(elated ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)
anova(model)
eta_squared(model)
r2(model)
model = lmer(anxious ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(sad ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(irritable ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(energetic ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)
model = lmer(angry ~ (time + I(time^2))*duration_above_250  + (1 | Id), data = light_metrics_current)
summary(model)

# KSS Model pre results
modelPre <- lmer(kss ~ time + I(time^2) + (1 | Id), 
                 data = current |> filter(season == "November-March"))
summary(modelPre)
coef_model <- fixef(modelPre)
b0 <- coef_model[1]         
b1 <- coef_model[2]
b2 <- coef_model[3]
x_min <- -b1 / (2 * b2)
y_min <- b0 + b1*x_min + b2*x_min^2
y_6am <- b0 + b1*6 + b2*6^2
cat("KSS value at 6am:", y_6am, "\n")
cat("Minimum KSS value (y_min):", y_min, "\n")
cat("Time at minimum KSS (x_min):", x_min, "\n")

# KSS Model post results
modelPost <- lmer(kss ~ time + I(time^2) + (1 | Id), 
                 data = current |> filter(season == "April-October"))
summary(modelPost)
coef_model <- fixef(modelPost)
b0 <- coef_model[1]         
b1 <- coef_model[2]
b2 <- coef_model[3]
x_min <- -b1 / (2 * b2)
y_min <- b0 + b1*x_min + b2*x_min^2
y_6am <- b0 + b1*6 + b2*6^2
cat("KSS value at 6am:", y_6am, "\n")
cat("Minimum KSS value (y_min):", y_min, "\n")
cat("Time at minimum KSS (x_min):", x_min, "\n")

# Model season interaction margins
model <- lmer(kss ~ time + I(time^2) + season + 
                time:season + I(time^2):season + 
                (1 | Id),
              data = current)
summary(model)
newdat <- data.frame(time = seq(4, 24, length.out = 200))
emm <- emmeans(
  model,
  ~ season | time,
  at = list(time = newdat$time)
)
emm_df <- as.data.frame(emm)

# Plot
plotKss <- ggplot(current, aes(x = time, y = kss, color = season)) +
  geom_point(alpha = 0.2) +
  geom_ribbon(
    data = emm_df,
    aes(x = time, ymin = lower.CL, ymax = upper.CL, fill = season),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  geom_line(
    data = emm_df,
    aes(x = time, y = emmean, color = season),
    linewidth = 1.2,
    inherit.aes = FALSE
  ) +
  scale_color_manual(values = c("November-March" = "#f08080", "April-October" = "#7fac7f")) +
  scale_fill_manual(values = c("November-March" = "#f08080", "April-October" = "#7fac7f")) +
  scale_x_continuous(limits = c(6, 24), breaks = seq(6, 24, by = 2)) +
  scale_y_continuous(limits = c(1, 10), breaks = seq(1, 10, by = 1)) +
  labs(
    x = "Time of Day",
    y = "KSS Score",
    color = "Months",
    fill = "Months"
  ) +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )
print(plotKss)


# Model light exposure interaction margins
model <- lmer(
  elated ~ (time + I(time^2)) * duration_above_250 +
    (1 | Id),
  data = light_metrics_current
)
summary(model)
exposure_values <- quantile(
  light_metrics_current$duration_above_250,
  probs = c(.25, .50, .75),
  na.rm = TRUE
)
time_values <- seq(4, 24, length.out = 200)
emm <- emmeans(
  model,
  ~ time | duration_above_250,
  at = list(
    time = time_values,
    duration_above_250 = unname(exposure_values)
  )
)
emm_df <- as.data.frame(emm)
emm_df$exposure_group <- factor(
  emm_df$duration_above_250,
  levels = unname(exposure_values),
  labels = c("Q1","Q2","Q3")
)

# Plot
plotElated <- ggplot() +
  geom_point(
    data = light_metrics_current,
    aes(x = time, y = elated),
    alpha = 0.15,
    color = "grey40",
    position = position_jitter(width = 0.05, height = 0)
  ) +
  geom_ribbon(
    data = emm_df,
    aes(
      x = time,
      ymin = lower.CL,
      ymax = upper.CL,
      fill = exposure_group,
      group = exposure_group
    ),
    alpha = 0.2
  ) +
  geom_line(
    data = emm_df,
    aes(
      x = time,
      y = emmean,
      color = exposure_group,
      group = exposure_group
    ),
    linewidth = 1.2
  ) +
  scale_x_continuous(
    limits = c(6, 24),
    breaks = seq(6, 24, by = 2)
  ) +
  scale_y_continuous(
    limits = c(1, 7), 
    breaks = seq(1, 7, by = 1)) +  
  scale_color_manual(
    values = c("#1F3A5F", "#3A7CA5", "#2CB1A6")
  ) +
  scale_fill_manual(
    values = c("#1F3A5F", "#3A7CA5", "#2CB1A6")
  ) +
  labs(
    x = "Time of Day",
    y = "Elatedness Score",
    color = "Light exposure",
    fill = "Light exposure"
  ) +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),
    panel.grid.minor = element_blank(),
    legend.position = "bottom"
  )
print(plotElated)

```

# Qualitative study results

```{r, message=FALSE, warning=FALSE}

# Theme coding table (17 participants)
themes_tbl <- tribble(
  ~row_id,
  ~health, ~child_development, ~positive_mood, ~cognitive_performance, ~work_motivation, 
  ~fatigue, ~alert_sleep_inertia, ~sleep_biological_clock, ~security, 
  ~morning_darkness, ~winter_effect, 
  ~morning_artificial_light_usage, ~loss_of_day_night_sense, 
  ~not_energy_save_more_electrical_light, ~evening_sunlight_not_important_late_work,
  ~wrong_meridian_selected_geographically,

   1,  0, 1, 1, 0, 1,  0, 1, 0, 0,  1, 0, 0, 0,  0, 0, 0,
   2,  0, 0, 0, 0, 0,  0, 0, 1, 0,  1, 0, 0, 0,  0, 0, 0,
   3,  0, 0, 0, 0, 0,  0, 0, 0, 1,  1, 1, 0, 0,  0, 0, 0,
   4,  0, 0, 0, 0, 0,  0, 0, 0, 0,  0, 0, 0, 0,  0, 0, 0,
   5,  0, 0, 0, 0, 0,  0, 0, 0, 0,  0, 0, 0, 0,  0, 0, 0,
   6,  0, 0, 1, 0, 0,  0, 0, 0, 0,  1, 1, 1, 1,  0, 1, 0,
   7,  0, 0, 1, 0, 0,  0, 0, 0, 1,  1, 0, 0, 0,  0, 0, 0,
   8,  0, 0, 0, 0, 0,  1, 1, 0, 0,  1, 1, 0, 0,  0, 0, 0,
   9,  0, 0, 0, 0, 0,  0, 0, 0, 0,  1, 0, 0, 0,  0, 0, 0,
  10,  0, 1, 0, 0, 0,  0, 0, 1, 0,  1, 0, 0, 1,  1, 0, 1,
  11,  0, 0, 1, 0, 0,  0, 0, 0, 0,  1, 0, 0, 0,  0, 0, 0,
  12,  0, 0, 0, 0, 0,  0, 1, 0, 0,  1, 0, 1, 1,  0, 1, 0,
  13,  1, 0, 0, 1, 0,  0, 0, 1, 0,  1, 0, 1, 1,  0, 0, 0,
  14,  0, 1, 0, 0, 0,  0, 0, 0, 0,  1, 0, 0, 0,  1, 0, 1,
  15,  0, 0, 0, 0, 0,  0, 0, 0, 0,  1, 0, 0, 0,  0, 0, 0,
  16,  0, 0, 1, 0, 0,  0, 0, 0, 0,  1, 0, 1, 0,  1, 0, 0,
  17,  0, 1, 1, 0, 0,  0, 1, 0, 0,  1, 0, 0, 1,  0, 0, 0
)

# Labels for negative effects
effect_labels <- c(
  health = "Reduced general health",
  child_development = "Concerns about child development",
  positive_mood = "Negative mood",
  cognitive_performance = "Reduced cognitive performance",
  work_motivation = "Reduced work motivation",
  fatigue = "Fatigue",
  alert_sleep_inertia = "Increased sleepiness / sleep inertia",
  sleep_biological_clock = "Sleep / circadian rhythm disruption",
  security = "Insecurity / fear"
)

# Labels for negative effects
reason_labels <- c(
  morning_darkness = "Morning darkness",
  winter_effect = "Winter morning darkness",
  morning_artificial_light_usage = "Increased use of electrical light in the morning",
  loss_of_day_night_sense = "Loss of day–night rhythm perception",
  not_energy_save_more_electrical_light = "No energy savings due to increased artificial lighting",
  evening_sunlight_not_important_late_work = "Evening daylight is less relevant due to late working hours",
  wrong_meridian_selected_geographically = "Misalignment of time zone with geographical location"
)

# Store variable names
effects <- names(effect_labels)
reasons <- names(reason_labels)

# Convert the effect variables to long format
effect_long <- themes_tbl |>
  select(row_id, all_of(effects)) |>
  pivot_longer(
    cols = -row_id,
    names_to = "effect",
    values_to = "effect_marked"
  ) |>
  filter(effect_marked == 1)

# Convert the reason variables to long format
reason_long <- themes_tbl |>
  select(row_id, all_of(reasons)) |>
  pivot_longer(
    cols = -row_id,
    names_to = "reason",
    values_to = "reason_marked"
  ) |>
  filter(reason_marked == 1)

# Create effect × reason co-occurrence table
heat_df <- effect_long |>
  inner_join(reason_long, by = "row_id") |>
  count(reason, effect, name = "freq") |>
  complete(
    reason = reasons,
    effect = effects,
    fill = list(freq = 0)
  ) |>
  mutate(
    percent = 100 * freq / nrow(themes_tbl),
    effect = factor(effect, levels = effects, labels = effect_labels),
    reason = factor(reason, levels = rev(reasons), labels = rev(reason_labels))
  )

# Heatmap plot
plotQual <-
  ggplot(heat_df, aes(x = effect, y = reason, fill = percent)) +
  geom_tile(color = "black", linewidth = 0.6) +
  geom_text(
    aes(label = ifelse(freq == 0, "", freq)),
    size = 3
  ) +
  scale_fill_gradient(
    low = "white",
    high = "#b2182b",
    name = "Frequency (%)"
  ) +
  scale_x_discrete(position = "top") +
  labs(
    x = "Negative effects",
    y = "Opinions about current time policy"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(
      angle = 30,
      hjust = 0,
      vjust = 0
    ),
    axis.ticks = element_blank()
  )
plotQual

# Combined panel
P_panel <- (
  (plotKss | plotElated) /
    free(plotQual, side = "l")
) +
  plot_layout(
    heights = c(1, 1.25)
  ) +
  plot_annotation(tag_levels = "A") &
  theme(
    axis.title.y = element_text(size = 12),
    plot.tag = element_text(face = "bold", size = 14)
  )
P_panel

# Save
ggsave(
  filename = "Figure-KSS Elated and qualitative.png", 
  plot = P_panel,
  width = 30,
  height = 30,
  units = "cm",
  dpi = 300
)

```


